---
title: "Linear models"
author: "Marta Coronado"
date: '2022-05-24'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
library(jtools)
library(ggplot2)
library(wesanderson)
library(car)
library(cowplot)
```

## Read data

From **Appendix 1** we take  the information on `repeat`, `location`, `LT50`, `Latitude`, `Longitude`, and `Strain` (Appendix 1, and Mortality data).

```{r info1, eval=FALSE}
info1 <- read.table("info1_revised.tab", header = T, sep = "\t", stringsAsFactors = T)
head(info1)
```

From **Appendix 2** we take the information on pollutants of each location, the distance to the closest city of >50k inhabitants, and the urban category.

```{r info2, eval=FALSE}
info2 <- read.table("info2_revised.tab", header = T, sep = "\t", stringsAsFactors = T)
info2
```

Data is merged by `Strain`, so we create a table with 97 observations and 16 variables.

```{r merge}
#data <- merge(info1, info2, by="Strain")
#write.table(data, "dataLM.tab", col.names = T, row.names = F, sep = "\t", quote = F)
data <- read.table("dataLM.tab", header = T, sep = "\t", stringsAsFactors = T)
head(data)
data$AreaClassification <- factor(data$AreaClassification, levels=c("Urban", "Semi-urban", "Rural"))
```

## Linear models

LM with all variables:

```{r lm-all}
lmAll = lm(LT50  ~ Longitude + Latitude +
           Copper + PM10 + PM2.5 + 
           Arsenic + Cadmium + Lead +
           AreaClassification,
         data = data) 
summary(lmAll)
```

### Backward stepwise regression
We want the model with only relevant variables. We can use `dropterm()` to remove variables that are not informative. The model should  be explaining the same amount of variance but only with the relevant variables.

The procedure is to remove variables 1 by 1, starting from the least significant, 

```{r lmAll-drop-term-1}
dropterm(lmAll, test = "F")
```

We first remove `Lead`, as it is the least significant variable.

```{r lm2}
lm2 = lm(LT50  ~ Longitude + Latitude +
           Copper + PM10 + PM2.5 + 
           Arsenic + Cadmium + 
           AreaClassification,
         data = data) 
summary(lm2) 


```


```{r lm-drop-term-2}
dropterm(lm2, test = "F")
```

We remove `PM10`.

```{r lm3}
lm3 = lm(LT50  ~ Longitude + Latitude +
           Copper + PM2.5 + 
           Arsenic + Cadmium + 
           AreaClassification,
         data = data)  
summary(lm3) 
```

```{r lm-drop-term-3}
dropterm(lm3, test = "F")
```

We remove `Cadmium`.

```{r lm4}
lm4 = lm(LT50  ~ Longitude + Latitude +
           Copper + PM2.5 + 
           Arsenic +  
           AreaClassification,
         data = data)  
summary(lm4) 
```

```{r lm-drop-term-4}
dropterm(lm4, test = "F")
```
We remove `Arsenic`.

```{r lm5}
lm5 = lm(LT50  ~ Longitude + Latitude +
           Copper + PM2.5 + 
           AreaClassification,
         data = data)  
summary(lm5) 

```


```{r lm-drop-term-5}
dropterm(lm5, test = "F")
```
We remove `PM2.5`.

```{r lm6}
lm6 = lm(LT50  ~ Longitude + Latitude +
           Copper + 
           AreaClassification,
         data = data)  
summary(lm6) 
```

```{r lm-drop-term-6}
dropterm(lm6, test = "F")
```

We remove `Longitude`.

```{r lm7}
lm7 = lm(LT50  ~  Latitude +
           Copper + 
           AreaClassification,
         data = data)  
summary(lm7) 
```

```{r lm-drop-term-7}
dropterm(lm7, test = "F")
```

We remove `Copper`.

```{r lm8}
lm8 = lm(LT50  ~  Latitude +
           AreaClassification,
         data = data)  
summary(lm8) 

Anova(lm8)
```

```{r lm-drop-term-8}
dropterm(lm8, test = "F")
```

Because now all terms are significant, it means that the relevant variables to explain LT50 are `Latitude` and `AreaClassification`.

We explain 12% of the variance, and the p-value is 0.008. The most important variable is  `Urban category - Urban`.  


### Plots


```{r plots}
plot_summs(lm8)

a<-effect_plot(lm8, pred = Latitude, interval = TRUE, plot.points = TRUE, 
            jitter = 0.05)

b<-effect_plot(lm8, pred = AreaClassification, interval = TRUE, plot.points = TRUE, 
            jitter = 0.05)
plot_grid(a,b,nrow=1,labels=c("A", "B"))
data$AreaClassification <- factor(data$AreaClassification, levels=c("Urban", "Semi-urban","Rural"))

data$map<-""
data[data$Strain=="ES_TEN" | data$Strain=="ES_COR" | data$Strain=="ES_GIM" | data$Strain=="ES_GUA" | data$Strain=="ES_TOM" | data$Strain=="IT_CAS" | data$Strain=="PT_REC",]$map<-"SE"
data[data$Strain=="AT_MUT" | data$Strain=="DE_MUN" | data$Strain=="DK_JUT" | data$Strain=="FI_AKA" | data$Strain=="SE_LUN" ,]$map<-"NE"
wilcox.test(data[data$map=="NE",]$LT50,data[data$map=="SE",]$LT50)
data$map <- factor(data$map, levels=c("SE", "NE"))

d<-ggplot(data, aes(x=AreaClassification, y=LT50, fill=AreaClassification)) + 
    geom_boxplot(show.legend = FALSE) + scale_fill_manual(values=c("gray25", "gray47", "gray77")) + geom_jitter(shape=16, position=position_jitter(0.2), show.legend = FALSE) + labs(x="", y = "LT50 values") + scale_x_discrete(breaks = c("Rural", "Semi-urban", "Urban"), labels = c("R", "SU", "U")) + 
  theme(axis.ticks = element_blank()) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))

c<-ggplot(data, aes(x=map, y=LT50, fill=map)) + 
    geom_boxplot(show.legend = FALSE) + scale_fill_manual(values=c("gray25", "gray47")) + geom_jitter(shape=16, position=position_jitter(0.2), show.legend = FALSE) + labs(x="", y = "LT50 values") + 
  theme(axis.ticks = element_blank()) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))

plot_grid(c,d)

 ggsave("Figure1CD.pdf", width = 7)
```

## Correlation variables

```{r corr}
corrplot(cor(unique(na.omit(data[,c("LT50","Longitude","Latitude","Copper","PM10","PM2.5","Arsenic","Cadmium","Lead")]))), method="number",  type="upper", tl.col="black",tl.srt=45)
```